Estimating associations between childhood exposure to non-persistent endocrine disruptors, corticosteroids, and neurodevelopment: A study based on the parametric g-formula

### 2023-11-22 output/paper/_si.qmd
Looked into ivreg for metabolites to chemicals.

```{r codebook-1}
#| label: supptbl-codebook-1
#| tbl-cap: Codebook for covariates used in RQ1

targets::tar_load(
  dplyr::contains("codebook_"),
  store = store
)
codebook_1
```

```{r codebook-2}
#| label: supptbl-codebook-2
#| tbl-cap: Codebook for covariates used in RQ2

codebook_2
```

```{r codebook-3}
#| label: supptbl-codebook-3
#| tbl-cap: Codebook for covariates used in RQ3

codebook_3
```

@@

```{r info-chems}
#| label: supptbl-info-chems
#| tbl-cap: Information about non-persistent EDCs

edcs_info <- myphd::edcs_information() |>
  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id)) |>
  tidylog::mutate(
    var = janitor::make_clean_names(
      short_name,
      replace = c(`-` = ""),
      case = "none"
    ) |>
      tolower()
  ) |>
  tidylog::relocate(var, .after = short_name)
gt::gt(edcs_info,
       groupname_col = "class") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    full_name = "Compound",
    short_name = "Symbol",
    var = "Variable name",
    pubchem_cid = "PubChem CID",
    parental_compound = "Parental compound"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  )
```

@@

```{r info-mets}
#| label: supptbl-info-mets
#| tbl-cap: Information about metabolites

mets_info <- readODS::read_ods(
  path = here::here("docs", "steroids.ods")
) |>
  tidylog::select(analyte, acronym, type, hmdb_id, cas_number)
gt::gt(mets_info,
       groupname_col = "type") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    analyte = "Metabolite",
    acronym = "Symbol",
    hmdb_id = "HMDB ID",
    cas_number = "CAS number"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  ) |>
  gt::tab_footnote(
    footnote = "Human Metabolome Database",
    locations = gt::cells_column_labels(
      columns = hmdb_id
    )
  ) |>
  gt::tab_footnote(
    footnote = "Chemical Abstracts Service",
    locations = gt::cells_column_labels(
      columns = cas_number
    )
  )
```

```{r lloq-mets}
#| label: supptbl-lloq-mets
#| tbl-cap: Lower limits of quantification for the corticosteroids

lloq_mets <- load_steroids()$loq |>
  tidylog::mutate(
    metabolite = stringr::str_replace(
      metabolite,
      "^X", ""
    )
  ) |>
  tidylog::rename(
    lloq = "loq"
  )
gt::gt(lloq_mets |> dplyr::arrange(metabolite)) |>
  gt::tab_footnote(
    footnote = "Lower limit of quantification expressed in nanograms per millilitre.",
    locations = gt::cells_column_labels(
      columns = lloq
    )
  )
```

@@

```{r weights-1sa}
#| label: supptbl-weights-1sa
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_1SA$dat_tbl,
  dat_plt = weights_1SA$dat_plt,
  sa_var = "e3_sex"
)$table
```

```{r weights-2sa}
#| label: supptbl-weights-2sa
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_2SA$dat_tbl,
  dat_plt = weights_2SA$dat_plt,
  sa_var = "e3_sex"
)$table
```

```{r weights-3sa}
#| label: supptbl-weights-3sa
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_3SA$dat_tbl,
  dat_plt = weights_3SA$dat_plt,
  sa_var = "e3_sex"
)$table
```

@@

```{r desc-chems-lod}
#| label: suppfig-desc-chems
#| fig-cap: Description of chemicals
#| fig-width: 6
#| fig-height: 7

viz_desc_vars(
  dat = load_dat_request()$dat,
  vars = vars_of_interest(append_to_chem = "cdesc")$chemicals,
  fct_levels = c(1, 2, 3, 4),
  is_chem = TRUE
)$plot
```

@@

```{r desc-mets-lod}
#| label: suppfig-desc-mets
#| fig-cap: Description of metabolites
#| fig-width: 6
#| fig-height: 7

viz_desc_vars(
  dat = myphd::extract_cohort(load_steroids()$desc,
                              id_var = id_var) |>
    tidylog::mutate(
      cohort = dplyr::case_when(
        cohort == "EDE" ~ "EDEN",
        cohort == "KAN" ~ "KANC",
        cohort == "MOB" ~ "MOBA",
        cohort == "RHE" ~ "RHEA",
        .default = cohort
      )
    ),
  vars = vars_of_interest()$metabolites |>
    paste0("_cdesc"),
  fct_levels = c(1, 2, 3, 4),
  is_chem = FALSE
)$plot
```

@@

```{r adrfs-rq3}
#| label: suppfig-adrfs-rq3
#| fig-cap: Average dose-response functions for RQ3
#| fig-width: 6
#| fig-height: 7

exposures_of_interest <- c(
  "cortisol production",
  "X11bHSD"
)
ret_adrfs <- lapply(
  exposures_of_interest, function(x) {
    plot_adrf(
      df_preds = marginal_gcomp_3 |>
        tidylog::filter(
          variable == x
        )
    )
  }
)
ret_adrfs[[1]]
#ret_adrfs[[2]]
```

@@

```{r marginal-1sa}
#| label: suppfig-marginal-1sa
#| fig-cap: Average comparisons for effect modification
#| fig-width: 6
#| fig-height: 7

tidy_res_meffects(
  df = marginal_comparisons_1SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-2sa}
#| label: suppfig-marginal-2sa
#| fig-cap: Average comparisons for effect modification
#| fig-width: 9
#| fig-height: 10

tidy_res_meffects(
  df = marginal_comparisons_2SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-3sa}
#| label: suppfig-marginal-3sa
#| fig-cap: Average comparisons for effect modification
#| fig-width: 5
#| fig-height: 5

tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

>>>>

```{r}
targets::tar_load(
  dplyr::contains("codebook_"),
  store = store
)
```

::: {#supptbl-codebook-1}
```{r codebook-1}
codebook_1
```
Codebook for covariates used in RQ1
:::

::: {#supptbl-codebook-2}
```{r codebook-2}
codebook_2
```
Codebook for covariates used in RQ2
:::

::: {#supptbl-codebook-3}
```{r codebook-3}
codebook_3
```
Codebook for covariates used in RQ3
:::

@@

::: {#supptbl-info-chems}
```{r info-chems}
edcs_info <- myphd::edcs_information() |>
  tidylog::select(-c(chem_id, smiles, ctd_id, exposome_explorer_id)) |>
  tidylog::mutate(
    var = janitor::make_clean_names(
      short_name,
      replace = c(`-` = ""),
      case = "none"
    ) |>
      tolower()
  ) |>
  tidylog::relocate(var, .after = short_name)
gt::gt(edcs_info,
       groupname_col = "class") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    full_name = "Compound",
    short_name = "Symbol",
    var = "Variable name",
    pubchem_cid = "PubChem CID",
    parental_compound = "Parental compound"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  )
```
Information about non-persistent EDCs
:::

@@

::: {#supptbl-info-mets}
```{r info-mets}
mets_info <- readODS::read_ods(
  path = here::here("docs", "steroids.ods")
) |>
  tidylog::select(analyte, acronym, type, hmdb_id, cas_number)
gt::gt(mets_info,
       groupname_col = "type") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    analyte = "Metabolite",
    acronym = "Symbol",
    hmdb_id = "HMDB ID",
    cas_number = "CAS number"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  ) |>
  gt::tab_footnote(
    footnote = "Human Metabolome Database",
    locations = gt::cells_column_labels(
      columns = hmdb_id
    )
  ) |>
  gt::tab_footnote(
    footnote = "Chemical Abstracts Service",
    locations = gt::cells_column_labels(
      columns = cas_number
    )
  )
```
Information about metabolites
:::

::: {#supptbl-lloq-mets}
```{r lloq-mets}
lloq_mets <- load_steroids()$loq |>
  tidylog::mutate(
    metabolite = stringr::str_replace(
      metabolite,
      "^X", ""
    )
  ) |>
  tidylog::rename(
    lloq = "loq"
  )
gt::gt(lloq_mets |> dplyr::arrange(metabolite)) |>
  gt::tab_footnote(
    footnote = "Lower limit of quantification expressed in nanograms per millilitre.",
    locations = gt::cells_column_labels(
      columns = lloq
    )
  )
```
Lower limits of quantification for the corticosteroids
:::

@@

::: {#supptbl-weights-1sa}
```{r weights-1sa}
tidy_res_weighted_fits(
  dat_tbl = weights_1SA$dat_tbl,
  dat_plt = weights_1SA$dat_plt,
  sa_var = "e3_sex"
)$table
```
Weights
:::

::: {#supptbl-weights-2sa}
```{r weights-2sa}
tidy_res_weighted_fits(
  dat_tbl = weights_2SA$dat_tbl,
  dat_plt = weights_2SA$dat_plt,
  sa_var = "e3_sex"
)$table
```
Weights
:::

::: {#supptbl-weights-3sa}
```{r weights-3sa}
tidy_res_weighted_fits(
  dat_tbl = weights_3SA$dat_tbl,
  dat_plt = weights_3SA$dat_plt,
  sa_var = "e3_sex"
)$table
```
Weights
:::

@@

::: {#suppfig-desc-chems}
```{r desc-chems-lod}
#| fig-width: 6
#| fig-height: 7

viz_desc_vars(
  dat = load_dat_request()$dat,
  vars = vars_of_interest(append_to_chem = "cdesc")$chemicals,
  fct_levels = c(1, 2, 3, 4),
  is_chem = TRUE
)$plot
```
Description of chemicals
:::

@@

::: {#suppfig-desc-mets}
```{r desc-mets-lod}
#| fig-width: 6
#| fig-height: 7

viz_desc_vars(
  dat = myphd::extract_cohort(load_steroids()$desc,
                              id_var = id_var) |>
    tidylog::mutate(
      cohort = dplyr::case_when(
        cohort == "EDE" ~ "EDEN",
        cohort == "KAN" ~ "KANC",
        cohort == "MOB" ~ "MOBA",
        cohort == "RHE" ~ "RHEA",
        .default = cohort
      )
    ),
  vars = vars_of_interest()$metabolites |>
    paste0("_cdesc"),
  fct_levels = c(1, 2, 3, 4),
  is_chem = FALSE
)$plot
```
Description of metabolites
:::

@@

::: {#suppfig-adrfs-rq3}
```{r adrfs-rq3}
#| fig-width: 6
#| fig-height: 7

exposures_of_interest <- c(
  "cortisol production",
  "X11bHSD"
)
ret_adrfs <- lapply(
  exposures_of_interest, function(x) {
    plot_adrf(
      df_preds = marginal_gcomp_3 |>
        tidylog::filter(
          variable == x
        )
    )
  }
)
ret_adrfs[[1]]
#ret_adrfs[[2]]
```
Average dose-response functions for RQ3
:::

@@

::: {#suppfig-marginal-1sa}
```{r marginal-1sa}
#| fig-width: 6
#| fig-height: 7

tidy_res_meffects(
  df = marginal_comparisons_1SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
Average comparisons for effect modification
:::

::: {#suppfig-marginal-2sa}
```{r marginal-2sa}
#| fig-width: 9
#| fig-height: 10

tidy_res_meffects(
  df = marginal_comparisons_2SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
Average comparisons for effect modification
:::

::: {#suppfig-marginal-3sa}
```{r marginal-3sa}
#| fig-width: 5
#| fig-height: 5

tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
Average comparisons for effect modification
:::

### 2023-11-22 output/paper/manuscript.qmd
Looked into ivreg for metabolites to chemicals.

format:
  docx:
    reference-doc: reference_document.docx
    citeproc: false
  # pdf:
  #   pdf-engine: lualatex
  #   number-sections: true
  #   toc: false
  #   keep-tex: false
  #   include-in-header:
  #     text: |
  #       \usepackage{lineno}
  #       \usepackage{rotating}
  #       \usepackage{float}
  #       \usepackage{typearea}
  #   include-before-body:
  #     text: |
  #       \linenumbers

@@

filters:
  - acronyms
  # - authors-block
  - ../../../../all_phd/_extensions/wordcount/citeproc.lua
  - ../../../../all_phd/_extensions/wordcount/wordcount.lua

@@

Children were assessed between December 2013 and February 2016, and included neurological testing and urine collection. Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment. Non-persistent \acr{edcs} assessed in urine samples from children included phthalate metabolites, phenols, and \acr{op} pesticide metabolites. A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems. The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].

@@

Levels of unprocessed non-persistent \acr{edcs} and corticosteroids are presented in @tbl-edc-desc and @tbl-met-desc, respectively.

>>>>

format:
  docx:
    reference-doc: reference_document.docx
    citeproc: true
  # pdf:
  #   pdf-engine: lualatex
  #   number-sections: true
  #   toc: false
  #   keep-tex: false
  #   include-in-header:
  #     text: |
  #       \usepackage{lineno}
  #       \usepackage{rotating}
  #       \usepackage{float}
  #       \usepackage{typearea}
  #   include-before-body:
  #     text: |
  #       \linenumbers

@@

filters:
  - acronyms
  # - authors-block
  # - ../../../../all_phd/_extensions/wordcount/citeproc.lua
  - ../../../../all_phd/_extensions/wordcount/wordcount.lua

@@

Children were assessed between December 2013 and February 2016, and included neurological testing and urine collection. Urine samples of the night before and the first morning void on the day of the visit were combined to provide a more reliable exposure assessment. Non-persistent \acr{edcs} assessed in urine samples from children included phthalate metabolites, phenols, and \acr{op} pesticide metabolites. A list of the environmental chemicals determined in urine samples and used for the present study is given in @supptbl-info-chems. Briefly, we analyzed a total of 7 phenols, 4 non-specific organophosphate pesticide metabolites, and 10 phthalate metabolites originating from 6 distinct phthalate parent compounds.
The laboratory protocols for the analysis are described elsewhere [@HaugSakhiCequier:2018].

@@

Levels of unprocessed non-persistent \acr{edcs} and corticosteroids, by cohort and overall, are presented in @tbl-edc-desc and @tbl-met-desc, respectively.

### 2023-11-22 output/presentations/cep_internal_231122/presentation.qmd
Looked into ivreg for metabolites to chemicals.

>>>>

## Even more difficulties

Most likely there is reverse causality (assumption of time-ordering from DAGs is not satisfied).

:::: {.columns}
::: {.column}
```{r}
coords <- tibble::tribble(
  ~name, ~x, ~y,
  "edc", 0, 0,
  "cortisol", 1, 0,
  "U", 0.5, 0.5
)
dag <- ggdag::dagify(
  cortisol ~ edc + U,
  edc ~ U,
  exposure = "edc",
  outcome = "cortisol",
  labels = c(
    "edc" = "EDC",
    "cortisol" = "cortisol",
    "U" = "confounders"
  ),
  coords = coords
)

ggdag::ggdag(
  dag,
  text = FALSE,
  use_labels = "label"
) +
  ggdag::theme_dag()
```
:::

::: {.column}
```{r}
coords <- tibble::tribble(
  ~name, ~x, ~y,
  "edc", 1, 0,
  "cortisol", 0, 0,
  "U", 0.5, 0.5,
  "iv", -0.5, 0
)
dag <- ggdag::dagify(
  cortisol ~ iv + U,
  edc ~ cortisol + U,
  exposure = "cortisol",
  outcome = "edc",
  labels = c(
    "edc" = "EDC",
    "cortisol" = "cortisol",
    "U" = "confounders",
    "iv" = "SNP"
  ),
  coords = coords
)

ggdag::ggdag(
  dag,
  text = FALSE,
  use_labels = "label"
) +
  ggdag::theme_dag()
```
:::
::::

* We cannot use 2-sample MR since there are no SNPs for the *outcome*.
* We do not have single metabolites but sums and ratios.
* I am not sure it makes sense...
